A Generalized Sphere Theorem for Positively Curved Combinatorial 3-Manifolds

### 2013-06-13 paper/paper.tex
updated split figures
bit of editing on abstract. worked on captions and other stuff for figures

\begin{abstract}
The generalized sphere theorem of Grove and Shiohama is a central result in the differential geometry of positively curved manifolds. It shows that if $M$ is a Riemannian manifold whose sectional curvature is everywhere greater than a positive constant and $M$ has diameter greater than half the maximum allowed by the Bonnet-Myers theorem, then $M$ must be homeomorphic to a sphere. In this paper, we present a novel discrete version of this theorem which applies to positively curved {\em combinatorial} 3-manifolds. That is, those for which at most five tetrahedra are incident at each edge. In the standard piecewise-linear metric, this bound is equivalent to requiring an angle deficit along each edge. Such spaces have been studied previously and found to satisfy a beautiful discrete version of the Bonnet-Myers theorem. Here we prove a corresponding {\em discrete} generalized sphere theorem using a complete census of positively curved 3-manifolds completed by Lutz and Sullivan.
\end{abstract}

@@

\begin{figure}
    \begin{center}
    {
	    \setlength\fboxsep{10pt}
		\setlength\fboxrule{0pt}
		\fbox{\includegraphics[width=0.75\linewidth]{figures/split_statistics.png}}
    }
    \begin{tabular} {| r | r | r | r | r | r | r |}
		\hline
		Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
		\hline
		\hline
		$1$&  &  &  &   &3    & 3    \\
		$H$&1 &  &  &10 &173  & 184  \\
		$2$&  &  &1 &7  &401  & 409  \\
		$J$&  &1 &1 &5  &2438 & 2445 \\
		$3$&  &  &  &   &1060 & 1060 \\
		$2H$&  &  &  &   &582  & 582  \\
		$4$&  &  &  &   &93   & 93   \\
		$2J$&  &  &  &   &11   & 11   \\
		\hline
		Total&1 &1 &2 &22   &4761   &4787  \\
		\hline
	\end{tabular}
    \end{center}
    \caption{Number of connected positively curved combinatorial 3-manifolds by topology and diameter, displayed graphically (above) and in table form (below)}
    \label{fig:type_statistics}
\end{figure}

\begin{figure}
    \begin{center}
    {
	    \setlength\fboxsep{10pt}
		\setlength\fboxrule{0pt}
		\fbox{\includegraphics[width=0.75\linewidth]{figures/edge_diameters.png}}
    }
	\begin{tabular} {| r | r | r | r | r | r | r |}	
		\hline
		Edge-Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
		\hline
		\hline
		$1$ &  &  &  &   &3    & 3    \\
		$2$ & 1&  &1 &17 &574  & 593  \\
		$3$ &  &1 &1 &5  &3498 & 3505 \\
		$4$ &  &  &  &   &675  & 675   \\
		$5$ &  &  &  &   &11   & 11   \\
		\hline
		Total&1 &1 &2 &22   &4761   &4787  \\
		\hline
	\end{tabular}
    \end{center}
    \caption{Number of connected positively curved combinatorial 3-manifolds by topology and edge-diameter, displayed graphically (above) and in table form (below)}
	\label{fig:type_statistics_edges}
\end{figure}

>>>>

\begin{abstract}
The generalized sphere theorem of Grove and Shiohama is an important result in the differential geometry of positively curved manifolds. It shows that if $M$ is a Riemannian manifold whose sectional curvature is everywhere greater than a positive constant and $M$ has diameter greater than half the maximum allowed by the Bonnet-Myers theorem, then $M$ must be homeomorphic to a sphere. In this paper, we present a novel discrete version of this theorem which applies to positively curved {\em combinatorial} 3-manifolds. That is, those with at most five tetrahedra incident at each edge, or equivalently, those with an angle deficit at each edge in the standard piecewise-linear metric. Such spaces have been studied previously and found to satisfy a beautiful discrete version of the Bonnet-Myers theorem. Here we prove the corresponding {\em discrete} generalized sphere theorem using a complete census of the positively curved 3-manifolds completed by Lutz and Sullivan.
\end{abstract}

@@

\begin{figure}
    \begin{center}
    {
	    \setlength\fboxsep{15pt}
		\setlength\fboxrule{0pt}
		\fbox{\includegraphics[width=0.9\linewidth]{figures/split_statistics.png}}
    }
    	\renewcommand{\arraystretch}{1.2}
    \begin{tabular} {| r | r | r | r | r | r | r |}
		\hline
		Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
		\hline
		\hline
		$1$&  &  &  &   &3    & 3    \\
		$H$&1 &  &  &10 &173  & 184  \\
		$2$&  &  &1 &7  &401  & 409  \\
		$J$&  &1 &1 &5  &2438 & 2445 \\
		$3$&  &  &  &   &1060 & 1060 \\
		$2H$&  &  &  &   &582  & 582  \\
		$4$&  &  &  &   &93   & 93   \\
		$2J$&  &  &  &   &11   & 11   \\
		\hline
		Total&1 &1 &2 &22   &4761   &4787  \\
		\hline
	\end{tabular}
    \end{center}
    \caption{Number of connected positively curved combinatorial 3-manifolds by topology and diameter, displayed graphically (above) and in table form (below). Note that all manifolds with diameter greater than half the maximum are 3-spheres. Results were computed using a 3-manifold census created by Lutz and Sullivan \cite{LutzSul_unpub}.}
    \label{fig:type_statistics}
\end{figure}

\begin{figure}
    \begin{center}
    {
	    \setlength\fboxsep{15pt}
		\setlength\fboxrule{0pt}
		\fbox{\includegraphics[width=0.9\linewidth]{figures/split_statistics.png}}
    }
    \renewcommand{\arraystretch}{1.2}
	\begin{tabular} {| r | r | r | r | r | r | r |}	
		\hline
		Edge-Diameter & $S^{3}/Q$ & $L(4,1)$ & $L(3,1)$ & $RP^{3}$ & $S^{3}$ & Total \\
		\hline
		\hline
		$1$ &  &  &  &   &3    & 3    \\
		$2$ & 1&  &1 &17 &574  & 593  \\
		$3$ &  &1 &1 &5  &3498 & 3505 \\
		$4$ &  &  &  &   &675  & 675   \\
		$5$ &  &  &  &   &11   & 11   \\
		\hline
		Total&1 &1 &2 &22   &4761   &4787  \\
		\hline
	\end{tabular}
    \end{center}
    \caption{Number of connected positively curved combinatorial 3-manifolds by topology and edge-diameter, displayed graphically (above) and in table form (below). Note that when diameters are measured in edges, there are manifolds with diameter greater than half the maximum which are {\em not} 3-spheres. Results were computed using a 3-manifold census created by Lutz and Sullivan \cite{LutzSul_unpub}.}
	\label{fig:type_statistics_edges}
\end{figure}
